---
title: "CrabCollabAll"
author: "Megsie Siple"
date: "9/7/2020"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(here)
library(Rmisc)
library(tidyverse)
library(ggsidekick) # just for plots: devtools::install_github("seananderson/ggsidekick")

source(here::here("Maia code","paramsScenarios.R")) # load params
source(here::here("Maia code","sizeMatrix.R")) # generate size transition matrix

crabdat <- read.csv(here::here('Heeia_crabdata_Apr2019.csv'))


```

## Simulations

The simulations are below; they are commented out because they take a while. To run the sims that are in the paper, un-comment and re-run.

### He'eia fishpond selectivity
```{r heeiasel}
source(here::here("Maia code","miscFunctions_heeia.R")) #
source(here::here("Maia code","runLeslieMC.R"))

runID = 'HeeiaSel_1000'
nsims.master = 1000
ts = format(Sys.time(), "%d%b_%Y") # for running
Name <- paste0(runID, '_', ts)
p <- proc.time()
test <- runLeslieMC(nsims.master <- nsims.master,
                    harvest.breaks = harvest.breaks)
proc.time() - p

writeOutMC(test, Name) ## write inits, lifehist, demography, stable harvest & risk table
## Generate summary text files (this happens instead of writing all raw data)
makeParStats(test, tc = 0, Name, write.file = T) ## default tc = 0
makeSADStats(test, Name = Name, write.file = T)
makeElastStats(test, Name = Name, write.file = T)
makeUsStats(test, Name = Name, write.file = T) # stationary harvest mortality

## quick inspection
params <- read.table(paste0(here::here('Maia code','outputs'),'/',Name,'_params.txt'),
                     header = T,sep = ",")
head(params)
# 
# dim(table(params$harvConst,params$tc))
```

#### Different selectivities ("Misty mountains")
```{r}
source(here::here("Maia code","miscFunctions.R")) #
source(here::here("Maia code","runLeslieMC.R"))

runID = 'proof_1000'
nsims.master = 1000
ts = format(Sys.time(), "%d%b_%Y")

Name <- paste0(runID, '_', ts)
p <- proc.time()
test <- runLeslieMC(nsims.master <- nsims.master,
                    harvest.breaks = harvest.breaks)
proc.time() - p

writeOutMC(test, Name) ## write inits, lifehist, demography, stable harvest & risk table
## Generate summary text files (this happens instead of writing all raw data)
makeParStats(test, tc = 0, Name, write.file = T) ## default tc = 0
makeSADStats(test, Name = Name, write.file = T)
makeElastStats(test, Name = Name, write.file = T)
makeUsStats(test, Name = Name, write.file = T) # stationary harvest mortality

## quick inspection
params = read.table(paste0(here::here('Maia code','outputs'),'/',Name,'_params.txt'),
                    header = T,sep = ",")
head(params)

dim(table(params$harvConst,params$tc))
```

## Figures

The figures in the paper, with the figure author in parentheses, are as follows:

1. Map of the fishpond with schematic (KALEO)
2. Size distribution of caught crabs (greyscale, with L95 etc) (MEGSIE)
3. CPUE over time (lines) (KALEO)
4. Total harvest per month (boxplots) (KALEO)
5. Weight vs. carapace width (KALEO)
6. Logistic selectivity plot (MEGSIE)
7. Heeia / turquoise distributions plot (MAIA)
8. "Misty mountains" plot with strict size cutoffs (MAIA)

### Figure 2
```{r, echo=FALSE, fig.width=8, fig.height=6,warning=FALSE}
all <- crabdat %>%
  filter(!Sex %in% c("FJ","J","NM")
         & !is.na(Sex)) %>%
  mutate(Sex = recode(Sex,`F` = "Female",M = "Male")) 
kept <- all %>%
  filter(Keep_Release_Dead=='Keep')

fig2 <- all %>%
  ggplot(aes(x=CWid_LR_mm)) + 
  geom_histogram(bins=30,alpha=.5) +
  geom_histogram(data=kept,aes(x=CWid_LR_mm)) +
  annotate('text',x=195,y=104,label = "paste(L[S95])",parse = TRUE,size = 5) +
  annotate('text',x=167.5,y=104,label = "paste(L[S50])",parse = TRUE,size = 5) +
  geom_segment(x=195,y=0,xend=195,yend=100,lty=2)+
  geom_segment(x=167.5,y=0,xend=167.5,yend=100,lty=2)+
  facet_wrap(~Sex) +
  ggsidekick::theme_sleek(base_size = 14) +
  xlab("Carapace width (mm)") +
  ylab("Count")

#pdf("Fig2.pdf",width = 8,height=4.5,useDingbats = F)
fig2
#dev.off()
```

### Figure 6
This figure shows the selectivity curve for the fishpond, and the size limits that we tested.

```{r}
Linf = 200
L95 = 195 #1st discontinuous break
keptcrabs <- crabdat %>% filter(Keep_Release_Dead=="Keep")
Lmin = min(keptcrabs$CWid_LR_mm,na.rm=T)
L50 = (L95+Lmin)/2
size.bins <- data.frame(matrix(c(11, 60, 61, 80, 81, 100, 101, 120, 121, 140,141, 160, 161, 180, 181, Linf, Linf + 1, 250), nrow = 2))
S <- vector()
L <- min(size.bins[1,]):max(size.bins[2,])

for(i in 1:length(L)){
  length.frac <- (L[i] - L50)/(L95-L50)
  S[i] <- 1 / (1+exp(-1*log(19) * length.frac))
}

plotdf <- data.frame(L=L,S=S)
linedf <- data.frame(L = apply(size.bins,MARGIN = 2,FUN = median)) %>% 
  mutate(S = 1/(1+exp(-1*log(19)*((L-L50)/(L95-L50)) ))  )

knife_edges <- data.frame(L=seq(min(size.bins[1,]),max(size.bins[2,]),1))%>%
  mutate(S1 = ifelse(L<76,0,1),
         S2 = ifelse(L<95.5,0,1),
         S3 = ifelse(L<152,0,1))
rulecolors <- RColorBrewer::brewer.pal(n = 4,name = "Blues")

Fig6 <- ggplot(plotdf,aes(x=L,y=S)) +
  geom_line() +
  geom_point(data=linedf,aes(x=L,y=S),colour="red") +
  geom_line(data=knife_edges,aes(x=L,y=S1),colour=rulecolors[2]) +
  geom_line(data=knife_edges,aes(x=L,y=S2),colour=rulecolors[3]) +
  geom_line(data=knife_edges,aes(x=L,y=S3),colour=rulecolors[4]) +  
  theme_classic(base_size = 14) +
  xlab("Length (mm)") +
  ylab("Fishery selectivity")

#pdf("Fig6.pdf",width = 6,height = 4,useDingbats = F)
Fig6
#dev.off()
```


### Figure 7
```{r}
runID <- 'HeeiaSel_1000'
ts <- '12May_2020'
Name <- paste0(runID, '_', ts)

# Load outputs
params <- read.table(paste0(here::here('Maia code','outputs'),'/',Name,'_params.txt'),
                     header = T,sep = ",")

dtp <-  params %>%
  select(rVal,harvConst,tc) %>%
  filter(tc==0) # doesn't matter which tc because they're all the same

#dtp
Fig7 <- dtp %>% 
  ggplot(aes(x=rVal)) +
  facet_wrap(~harvConst,scales="free_y") +
  scale_fill_brewer('Minimum \ncapture \nsize (mm)') +
  geom_density(alpha = 0.5,fill='turquoise') +
  xlim(c(-2,1)) +
  xlab('Population growth rate (r)') +
  ylab('Density') +
  theme_sleek() +
  geom_vline(xintercept = 0,lty=1)

#  pdf("Fig7_R2.pdf",width = 10,height = 5, useDingbats = FALSE)
Fig7
#  dev.off()  
```


### Figure 8
This figure shows the fixed size limits we tested, including the state size limit. 

```{r}
Name <- "proof_1000_07Oct_2020"
params <- read.table(paste0(here::here('Maia code','outputs'),'/',Name,'_params.txt'),
                     header = T,sep = ",")
head(params)
# Check mean rval at each harvest rate
( medians <- params %>%
    group_by(harvConst,tc) %>%
    summarize(medR = median(rVal)) )

Fig8 <- params %>%
  select(rVal,harvConst,tc) %>%
  ggplot(aes(x=rVal,fill=factor(tc))) +
  xlim(c(-2,1)) +
  facet_wrap(~harvConst) +
  scale_fill_brewer('Minimum \ncapture \nsize (mm)') +
  geom_density(alpha = 0.5) +
  geom_vline(data=medians, aes(xintercept=medR),colour='darkgrey',lty=2) +
  xlab('Population growth rate (r)') +
  ylab('Density') +
  theme_sleek() +
  geom_vline(xintercept = 0,lty=1)

# tiff("Fig3_1000_GenSel.tiff",width = 8,height = 5,units = 'in',res=200)
Fig8
# dev.off()
```


### Table 2

```{r}
Name <- "proof_1000_07Oct_2020"  # for reading
params <- read.table(paste0(here::here('Maia code','outputs'),'/',Name,'_params.txt'),
                     header = T,sep = ",")
doublingTime <- mean(params$tDouble) 
statsUsMC <- read.table(paste0(here::here('Maia code','outputs'),'/',Name,'_StatsUsMC.txt'),
                        header = T,sep = ",")
x <- params %>%
  filter(harvConst==0) %>%
  group_by(tc) %>%
  summarize(meanR = mean(rVal),
            meangenTime = mean(genTime),
            meanR0 = mean(Rzero))

unfishedPopGrowthRate <- x$meanR[1]
minStationaryHarvMort <- statsUsMC$minVal[1]
meanPerGenGrowthRate <- x$meanR0[1]
meanGenTime <- x$meangenTime[1]

Table2 <- data.frame( Parameter = c("population doubling time",
                                    "unfished pop growth rate",
                                    "min stationary harvest mortality",
                                    "mean per-gen growth rate",
                                    "mean generation time"),
                      `Value from simulation` = c(doublingTime,
                                                  unfishedPopGrowthRate,
                                                  minStationaryHarvMort,
                                                  meanPerGenGrowthRate,
                                                  meanGenTime))
Table2
```

### In-text values
Normally I would write the text with these variables just effortlessly embedded in the text, but these will have to do. Double check that the `params` object is the one you want before running. 

```{r}
# Different harvest rates (blue mountains)
Name <- "KnifeEdge_1000_07Feb2020"  # for reading
params <- read.table(paste0(here::here('Maia code','outputs'),'/',Name,'_params.txt'),
                     header = T,sep = ",")

# what is the mean growth rate at each harvConst for the heeia selectivity case?
params %>%
  group_by(harvConst) %>%
  summarize(meanR = mean(rVal),
            meangenTime = mean(genTime),
            meanR0 = mean(Rzero)) 

# what is the pop doubling time?
mean(params$tDouble) 

# how many combos of harvConst and min cap size?
length(table(unique(params[,c('tc','harvConst')])))

# which harvest rate produces no negative growth rates?
params %>%
  group_by(harvConst,tc) %>%
  summarize(min(rVal)) %>%
  as.data.frame()

# what is the minimum size captured in the data?
76

# what is the min tagged size (for putting survivorship from CJS into the matrix model)
crabdat %>%
  filter(Tag_YNNew =="Yes") %>%
  summarize(min.tagged.size = min(CWid_LR_mm))

# What is the avg min survivorship?
outUs <- read.table(paste0(here::here('Maia code','outputs'),'/',Name,'_outUS.txt'),
                    header = T,sep = ",")

outUs %>%
  group_by(tc) %>%
  summarize(minAvgSurv = min(Us))

```
