---
title: "Model methods and results"
author: "M. Siple"
date: "2/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
crabdat <- read.csv(here::here('Heeia_crabdata_Apr2019.csv'))

```

## Methods


### Estimate of adult survival

Estimates of adult survival ($\phi_a$) and recapture probability ($p_a$) were estimated from mark-recapture data using a Cormack-Jolly-Seber (CJS) model [Cormack (1964); Jolly:1965; Seber:1965]. The CJS model assumes that the population is open and that the population size changes over time due to natural mortality. Mark and recapture data were divided into monthly sampling periods, such that every sampling month following the initial tagging (N = 112 individuals) was one time period. We assumed that parameters were constant, i.e., that survival probability did not change through time. Priors for each parameter were uniform between 0 and 1.

The model was sampled from four chains with 1,000 post-warmup draws per chain, thinning every other iteration. The CJS model was implemented in Stan (cite).


### Estimate of fishery selectivity

Hawai‘i state size restrictions for Samoan crab are XXX. In He'eia fishpond, the control rule used is that only males of carapace width YYY are kept. We estimated fishery selectivity from kept crabs surveyed during the mark-recapture experiment, using length compositions as in Nadon 2017 (cite). 

$$
S = \frac{1}{1+exp(-ln(19) \frac{L-L_{S50}}{L_{S95-LS10}})}
$$
where $L$ is the length at which selectivity is being estimated, $L_{S50}$ is the length at 50% selectivity, and $L_{S95}$ is the length at 95% selectivity. Following Nadon (2017), we used the first discontinuous break in size composition of kept male crabs as the size at 95% selectivity ($L_{S95}$= 169 mm), and the midpoint between the minimum caught size () and $L_{S95}$ as the estimate of $L_{S50}$.

```{r, echo=FALSE, fig.width=8, fig.height=6,warning=FALSE}
all <- crabdat %>%
  filter(!Sex %in% c("FJ","J","NM")
         & !is.na(Sex)) %>%
  mutate(Sex = recode(Sex,`F` = "Female",M = "Male")) 
kept <- all %>%
  filter(Keep_Release_Dead=='Keep')

fig2 <- all %>%
  ggplot(aes(x=CWid_LR_mm)) + 
  geom_histogram(bins=30,alpha=.5) +
  geom_histogram(data=kept,aes(x=CWid_LR_mm)) +
  annotate('text',x=195,y=104,label = "paste(L[S95])",parse = TRUE,size = 5) +
  annotate('text',x=167.5,y=104,label = "paste(L[S50])",parse = TRUE,size = 5) +
  geom_segment(x=195,y=0,xend=195,yend=100,lty=2)+
  geom_segment(x=167.5,y=0,xend=167.5,yend=100,lty=2)+
  facet_wrap(~Sex) +
  ggsidekick::theme_sleek(base_size = 14) +
  xlab("Carapace width (mm)") +
  ylab("Count")

pdf("Fig2.pdf",width = 8,height=4.5,useDingbats = F)
fig2
dev.off()
```
Selectivity for each size class is as follows:
```{r, echo=F}
Linf = 200
L95 = 195 #1st discontinuous break
keptcrabs <- crabdat %>% filter(Keep_Release_Dead=="Keep")
Lmin = min(keptcrabs$CWid_LR_mm,na.rm=T)
L50 = (L95+Lmin)/2
size.bins <- data.frame(matrix(c(11, 60, 61, 80, 81, 100, 101, 120, 121, 140,141, 160, 161, 180, 181, Linf, Linf + 1, 250), nrow = 2))
S <- vector()
L <- min(size.bins[1,]):max(size.bins[2,])

for(i in 1:length(L)){
  length.frac <- (L[i] - L50)/(L95-L50)
  S[i] <- 1 / (1+exp(-1*log(19) * length.frac))
}

plotdf <- data.frame(L=L,S=S)
linedf <- data.frame(L = apply(size.bins,MARGIN = 2,FUN = median)) %>% 
  mutate(S = 1/(1+exp(-1*log(19)*((L-L50)/(L95-L50)) ))  )

knife_edges <- data.frame(L=seq(min(size.bins[1,]),max(size.bins[2,]),1))%>%
                      mutate(S1 = ifelse(L<76,0,1),
                             S2 = ifelse(L<95.5,0,1),
                             S3 = ifelse(L<152,0,1))
rulecolors <- RColorBrewer::brewer.pal(n = 4,name = "Blues")

Fig6 <- ggplot(plotdf,aes(x=L,y=S)) +
        geom_line() +
        #geom_vline(data=linedf,aes(xintercept = L),colour="red",lty=2) +
        geom_point(data=linedf,aes(x=L,y=S),colour="red") +
        geom_line(data=knife_edges,aes(x=L,y=S1),colour=rulecolors[2]) +
        geom_line(data=knife_edges,aes(x=L,y=S2),colour=rulecolors[3]) +
        geom_line(data=knife_edges,aes(x=L,y=S3),colour=rulecolors[4]) +  
        theme_classic(base_size = 14) +
        xlab("Length (mm)") +
        ylab("Fishery selectivity")
 
pdf("Fig6.pdf",width = 6,height = 4,useDingbats = F)
Fig6
dev.off()
```
#### Minimum size at capture in this study
```{r}
crabdat %>% summarize(min(CWid_LR_mm,na.rm=T))
```

### Stage-based simulation model

The simulation model used to test different fishery scenarios is a size-structured model with natural mortality estimated from crabs in He'eia fishpond and recruitment parameters based on literature values and estimates from the size structure of male and female crabs from the mark-recapture experiment. [see Maia doc LeslieMK_mcs.R and Maia's written supplement for the rest of the methods].

### Population model
#### Size-transition matrix
Crabs were divided into 7 size bins, based on the minimum and maximum carapace width found in the survey:

```{r}
size.bins <- c("60-80","80-100","100-120","120-140","14-160","160-180","180-200")

```

## Results

### Estimate of adult survival
The parameter $\phi_a$ is a survival rate per time period represented in the mark-recapture data. We estimated a survival rate of 0.966 [95% ci 0.889, 0.999] between each sampling period. This corresponds to a monthly survival rate of 0.947 or $M_{adult} = 0.054$ for the monthly time steps in the simulation model. We estimated a recapture probability of $p_a$ = 0.014 [0.011, 0.024].  