---
title: "Leslie Matrix Simulation: Samoan Crab"
author: "M Kapur adapted from A Aires da Silva"
date: "2017 April - 2020 January"
output:
  html_notebook
---
The final objective of this code is to compute a general stochastic demographic analysis for Samoan crab. The original file was written by Maia Sosa Kapur, and has been recently edited by Megsie Siple to apply to Samoan crab. This mostly involves parameterizing. Adult survivorship is separately estimated from mark-recapture data. See Methods.Rmd for a description of that part.

Parameters that are currently simulated are:

1. EPR, the number of spawning events per year (runif(1,0,3)) - ACTUALLY FIXED @ 2
2. longevDraw (maximum age) - rnorm(1, mean = 16, sd = 2) - ACTUALLY FIXED @ 16
3. SurvAge (vector of survivorship @ age) - runif(longevDraw, 0.2, 0.35) takes on length of longevDraw & coerced to decrease


```{r runLeslieMC, message = F, warning = F, eval = F}
library(ggplot2)
library(here)
library(Rmisc)
library(dplyr)
library(tidyr)

source(here::here("Maia code","paramsScenarios.R"))
source(here::here("Maia code","sizeMatrix.R")) ## generate the size matrix separately
source(here::here("Maia code","miscFunctions.R"))

## FUNCTION runLeslieMC
## Purpose: create Leslie matrix, applying simulations described above.
runLeslieMC = function(nsims.master, harvest.breaks) {
  # Define output table of MC simulations for demographic parameters
  params.to.save = c('eigenDomAH',
                     'rVal',
                     'tDouble',
                     'Rzero',
                     'genTime',
                     'harvConst',
                     'tc')
  OutParams = data.frame(matrix(
    NA,
    nrow = nsims.master * harvest.breaks * length(tc.vector),
    ncol = length(params.to.save)
  ))
  names(OutParams) = params.to.save
  
  NamesLifeHist <- names(size.bins)
  InLifeHistMC = matrix(NA,
    nrow = nsims.master,
    ncol = length(NamesLifeHist) * 2 + 1,
    dimnames = list(NULL, c(
      "simID",
      paste('Fec', NamesLifeHist),
      paste('Surv', NamesLifeHist)
    ))
  )
  
  
  # Define output table of MC simulations for SAD
  OutSADMC = data.frame(matrix(
    NA,
    nrow = nsims.master * harvest.breaks,
    ncol = ncol(size.bins) + 1
  ))
  names(OutSADMC) = c(paste0(colnames(size.bins)), 'harvConst')
  
  # Define output table of MC simulations for elasticities
  NamesElast  = c(paste0('elastFec', colnames(size.bins)),
                  paste0('elastSurv', colnames(size.bins)),
                  'elastCheck')
  OutElastMC = data.frame(matrix(
    NA,
    nrow = nsims.master * length(h.vector) * length(tc.vector),
    ncol = length(NamesElast) + 1
  ))
  names(OutElastMC) =  c(paste0(NamesElast), 'harvConst')
  
  # Define output table of MC simulations for Us (sustainable harvest rate)
  OutUsMC <-
    data.frame(matrix(NA, nrow = nsims.master * length(tc.vector), 3))
  names(OutUsMC) = c('simID', 'Us', 'tc')
  
  # MONTE CARLO simulations
  for (isim in 1:nsims.master) {
    if (isim %% 100 == 0) {
      cat("isim ",isim,"\n")
    } ## report every 100 sims

    ## Run Pop Model
    longevDraw = runif(1,7,16) ## lit reported range
    
    # mh = hoenig(
    #   hoenig.slope = hoenig.slope,
    #   hoenig.int = hoenig.int,
    #   longevDraw = longevDraw
    # )
    # 
    # nat.survivorship.temp = nat.mort(
    #   longevDraw = longevDraw,
    #   zeta = 0.3,
    #   mh = mh,
    #   age.est.vec = 1:longevDraw
    # )[, 'SURVIVORSHIP'] ## values from Oneill; age based matrix
    
    # longevDraw = 16
    # Natural Mortality: Randomly choose survivorship from uniform distribution, dims of longevity (which varies)
    SurvAge <- sort(runif(longevDraw, 0.01, 0.04), decreasing = F)

    # cat(SurvAge,"\n")
    # survPlot(SurvAge, size.matrix)
    ## N is placeholder array of 1
    N <- array(1, dim = c(length(SurvAge), ncol(size.matrix)))
    ## Multiply by STM - this becomes diagonal of Lefkovich
    
    surv.vector <-  (N * SurvAge) %*% size.matrix %>%
      round(digits = 4) %>%
      apply(.,2,mean) %>%
      t() %>%
      as.data.frame()
    surv.vector[1] <- 0.064 ## coerce to this value
    colnames(surv.vector) <- NULL

    ## Fecundity
    FX = FX.func(size.bins) ## just expected egg output
    # EPR = ceiling(abs(rnorm(1,1,1)))
    # cat(EPR,"\n")
    EPR = 2
    fecundity.vector = FX * SR * EPR ## per capita female birth rate
    # cat(fecundity.vector,"\n")
  
    #   age.est.vec = estimate.age(
    #   size.bins = size.bins,
    #   Linf = Linf,
    #   K = K,
    #   tZero = tZero,
    #   longevDraw = longevDraw
    # )
    # # 

    # # 
    

    # surv.vector = nat.survivorship.temp %*% size.matrix ## multiply by size matrix to get surviving portion within each size class; see Punt 2003. These are decoupled from harvest
    
    ## GENERATE VECTORS FOR ESTIMATED RECRUITS
    # %>% apply(., 1, prop.table) %>% round(digits = 4)
    #### LESLIE MATRIX ####
    LeslieMat = matrix(0,
                       nrow = length(surv.vector),
                       ncol = length(surv.vector))
    # Populate first row with fertility vector
    LeslieMat[1, ] = fecundity.vector
    # Populate the rest of the matrix with survival terms (Pi) - only on diagonal
    for (i in 2:length(surv.vector)) {
      j = i - 1
      LeslieMat[i, j] = surv.vector[,i - 1]
    }
    
    # Populate LH list with draws for each run
    InLifeHistMC[isim,] = c(isim,fecundity.vector,as.numeric(surv.vector))


    ## Estimate the STATIONARY HARVEST (Us) using the BISSECTION METHOD
    ## [only relies on Leslie Mat, which already has nat mort in place]
    for (t in 1:length(tc.vector)) {
      ## all ages, only L50, only legal size
      indexUs = (isim - 1) * length(tc.vector) + t  ## for OutUS
      
      Us = runBissect(Ulow,
                      Uhigh,
                      bissectIters,
                      bissectConv,
                      LeslieMat,
                      tc = tc.vector[t]) ## generate different Us estimates for each TC value for each sim of Leslie
      OutUsMC[indexUs,] = as.vector(c(isim, Us, tc.vector[t]))
    }
    
    
    ## LEGAL ONLY BISSECTION
    ## Estimate the STATIONARY HARVEST (Us) using the BISSECTION METHOD [only relies on Leslie Mat, indifferent to harvest rate]
    # Us = runBissect(Ulow,
    #                 Uhigh,
    #                 bissectIters,
    #                 bissectConv,
    #                 LeslieMat,
    #                 tc = tc.vector[t]) ## generate different Us estimates for each TC value for each sim of Leslie
    # OutUsMC[isim, ] = as.vector(c(isim, Us, tc.vector[t]))
    
    
    
    #### HARVEST MATRIX ####
    ## Use Decrement -- we want to apply the entire suite of harvest scenarios to a single leslie
    
    for (h in 1:length(h.vector)) {
      for (t in 1:length(tc.vector)) {
        ## all ages, only L50, only legal size
        # t = 4 ## hard assignment of tc vector for legal sizes
        HarvestMat =  makeHarvestMat(size.bins, harvConst = h.vector[h], tc = tc.vector[t])
        # index = (isim-1)*length(h.vector) + h ## original just with H
        index = (isim - 1) * length(h.vector) * length(tc.vector) + (h - 1) *
          length(tc.vector) + t ## for bigger MCs
        
        #### DEMOGRAPHIC PARAMETERS ####
        # Dominant EIGENVALUE of the AH matrix
        AHmat = LeslieMat %*% HarvestMat ## MATRIX MULTIPLICATION OPERATOR
        # cat(diag(AHmat),"\n")
        eigenDomAH = max(abs(eigen(AHmat)$values)) ## a value of zero means it is NOT invertible...
        
        # Instantaneous rate of population increase (r) [exp(r) = lambda]
        rVal = log(eigenDomAH)
        # if (h > 8) {cat(rVal,"\n")}
        # Population doubling time (t2) and halving time (t0.5)
        tDouble = log(2) / rVal
        tHalf = log(.5) / rVal
        
        # NET REPRODUCTIVE RATE, R0 (see Caswell 2001, p. 126)
        ## Decompose the AH matrix (see Caswell 2001, p. 110, eq. 5.1)
        # Compute the T matrix (describing transitions)
        Tmat = AHmat
        Tmat[1,] = 0
        # Compute the F matrix (describing reproduction)
        Fmat = matrix(0, ncol(LeslieMat), ncol(LeslieMat))
        Fmat[1,] = AHmat[1,]
        ## Compute the FUNDAMENTAL MATRIX (N) (see Caswell 2001, p. 112, eq. 5.7)
        # Compute the IDENTITY MATRIX (I)
        Imat = matrix(0, ncol(LeslieMat), ncol(LeslieMat))
        diag(Imat) = 1
        # Compute N matrix
        Nmat = solve(Imat - Tmat)
        ## Compute R MATRIX (R) (see Caswell 2001, p. 126, eq. 5.64)
        Rmat = Fmat %*% Nmat
        ## Compute NET REPRODUCTIVE RATE, R0 (see Caswell 2001, p. 126)
        Rzero = max(abs(eigen(Rmat)$values))
        
        # GENERATION TIME, Abar (see Caswell 2001, p. 128)
        # How to compute Abar? Use T meanwhile...
        genTime = log(Rzero) / log(eigenDomAH)
        
        # Populate vector with output demographic parameters
        # OutParams[isim, ] = c(eigenDomAH, rVal, tDouble, Rzero, genTime, h)
        OutParams[index, ] = c(eigenDomAH,
                               rVal,
                               tDouble,
                               Rzero,
                               genTime,
                               h.vector[h],
                               tc.vector[t])
        
        
        ### FIX PROBLEM - hitting the upper bound for high tc values!
        
        ## ELASTICITY ANALYSIS (see Caswell 2001, p. 128)
        # Compute STABLE AGE DISTRIBUTION, SAD (w)
        w = abs(eigen(AHmat)$vectors[, 1])
        # Normalize w
        w = w / sum(w)
        # Compute REPRODUCTIVE VALUE vector (v)
        v = abs(eigen(t(AHmat))$vectors[, 1])
        # Compute inner product <w,v>
        iProd = sum(w * v)
        
        ## Compute SENSITIVITY matrix (see Caswell 2001, p. 209, eq. 9.12)
        # Define dimensions of sensitivity matrix
        SensMat = matrix(0, ncol(size.bins), ncol(size.bins))
        # Populate matrix with sensitivity elements
        for (i in 1:ncol(size.bins)) {
          for (j in 1:ncol(size.bins))
            SensMat[i, j] = (v[i] * w[j]) / iProd
        }
        
        ## Compute ELASTICITY matrix (see Caswell 2001, p. 226, eq. 9.70)
        # Define dimensions of elasticity matrix
        ElastMat = matrix(0, ncol(size.bins), ncol(size.bins))
        # Populate matrix with sensitivity elements
        for (i in 1:ncol(size.bins)) {
          for (j in 1:ncol(size.bins))
            ##! this is wrt lambda - does it matter?
            ElastMat[i, j] = (AHmat[i, j] / eigenDomAH) * SensMat[i, j]
        }
        ## Compute FECUNDITY and SURVIVAL ELASTICITIES for age-groups
        # Extract first row (fecundity elasticities) from "ElastMat"
        ElastFec = ElastMat[1,]
        # Extract vector with survival elasticities from "ElastMat"
        ElastSurv = rep(0, ncol(size.bins))
        for (i in 2:ncol(size.bins)) {
          j = i - 1
          ElastSurv[i - 1] = ElastMat[i, j]
        }
        # meaningless names just placeholders
        elastFec_1 = ElastFec[1]
        elastFec_2 = ElastFec[2]
        elastFec_3 = ElastFec[3]
        elastFec_4 = ElastFec[4]
        elastFec_5 = ElastFec[5]
        elastFec_6 = ElastFec[6]
        elastFec_7 = ElastFec[7]
        
        elastSurv_1 = ElastSurv[1]
        elastSurv_2 = ElastSurv[2]
        elastSurv_3 = ElastSurv[3]
        elastSurv_4 = ElastSurv[4]
        elastSurv_5 = ElastSurv[5]
        elastSurv_6 = ElastSurv[6]
        elastSurv_7 = ElastSurv[7]
        
        elast.vals = c(
          elastFec_1,
          elastFec_2,
          elastFec_3,
          elastFec_4,
          elastFec_5,
          elastFec_6,
          elastFec_7,
          elastSurv_1,
          elastSurv_2,
          elastSurv_3,
          elastSurv_4,
          elastSurv_5,
          elastSurv_6,
          elastSurv_7
        )
        # Check on elasticity computations
        elastCheck = sum(ElastFec) + sum(ElastSurv)
        
        ## Populate output table of elasticity analysis
        OutElast = c(elast.vals, elastCheck)
        
        ## Populate output tables
        # Table of SAD
        OutSAD = rep(NA, length(OutSADMC))
        OutSAD = w
        OutSAD[ncol(size.bins)] = sum(w[ncol(size.bins) - 1:ncol(size.bins)])
        
        OutSADMC[index,] = as.vector(c(OutSAD, h.vector[h]))
        OutElastMC[index,] = as.vector(c(OutElast, h.vector[h]))
      }
    }
  }
  
  return(
    list(
      'params' = OutParams,
      'lifehist' = InLifeHistMC,
      'inits' = inits,
      'outSAD' = OutSADMC,
      'outElast' = OutElastMC,
      'outUs' = OutUsMC
    )
  )
  }

```


Run & Plot outcomes of Leslie matrix simulation.
```{r, eval = F}
## ON GUSTAV: c(nsims, harvest.breaks, tc)
## 18 hours (10K,10,4)
## 2.45 hours (10K,10,0)
## 1 HOUR (10K,5,0)
## 10 Mins(1000,10,0)
## 14 hours(1 million, 0,0)
## 8.5 mins(100k, 0,0)

## On V003
## < 10 hours (10K,10,4)

runID = 'newMort'
nsims.master = 100
ts = format(Sys.time(), "%d%b%Y")
Name <- paste0(runID, '_', ts)
p = proc.time()
test = runLeslieMC(nsims.master = nsims.master,
                   harvest.breaks = harvest.breaks)
proc.time() - p

writeOutMC(test, Name) ## write inits, lifehist, demography, stable harvest & risk table
## Generate summary text files (this happens instead of writing all raw data)
makeParStats(test, tc = 0, Name, write.file = T) ## default tc = 0
makeSADStats(test, Name = Name, write.file = T)
makeElastStats(test, Name = Name, write.file = T)
makeUsStats(test, Name = Name, write.file = T)

## quick inspection
params = read.table(paste0(here::here("Maia code","outputs"),"/",Name,"_params.txt"),
                    header = T,sep = ",")
with(subset(params, tc == 0 & harvConst == 0.9), hist(rVal))
# plot(1,type = 'n', xlim = c(-3,3),ylim = c(0,10), yaxt = 'n', ylab = 'mk')
# d = with(subset(params, tc == 0 & harvConst == 0.9), density(rVal))
# polygon(d, col = "grey", border = 'white')

harvestPlot(OutLeslieMC.FILE = params, Name = Name, 
form = 'png') ## save r values vs h,tc plot
graphics.off()
```


Read life history info, summarize survival and fecundity by size class.
```{r, echo = TRUE}
lifehist = read.table(paste0(here::here("Maia code","outputs"),"/",Name,"_lifehist.txt"), header = T, sep = ',')
    # lifehist %>% 
    #   select(c(Surv.11.58,Surv.59.65,Surv.66.80,Surv.81.90,Surv.91.102,Surv.103.115,Surv.116.)) %>%
    #   summarise_all(funs(mean))

#MCS: funs() is soft-depracated so here is a new way to summarize
lifehist %>%
  gather(HEADER,simID) %>% 
  group_by(HEADER) %>%
  summarize(meanval = mean(simID))
## inspect the outputs for gross errors
# head(test$lifehist)

param.plots(test) ## inspect demographic plots
```


Save plots in files
```{r}
savePlots(param.plots, OutLeslieMC = OutLeslieMC, Name = Name, vals = 'params', form = 'jpg', height = 1, width = 10) ## save demographic plots
# harvestPlot(OutLeslieMC.FILE = test$params, Name = Name, form = 'png') ## save r values vs h,tc plot
savePlots(plotElast, test, Name, vals = 'elast', form = 'png', height = 550, width = 725) ## save elasticity plot

graphics.off()

harvestPlot2(OutLeslieMC.FILE = params, Name = Name, 
             form = 'png', bgcol = c('black',"white")[1]) ## save r values vs h,tc plot
graphics.off()

# View(makeParStats(test, tc = 0, Name, write.file = F))
# View(makeSADStats(test, Name, write.file = F))
# View(makeElastStats(test, Name, write.file = F))
# View(makeUsStats(test, Name, write.file = T))


# params = read.table('G:/KONA CRAB/outputs/MasterRun_0515_15May2017_params.txt', header = T, sep = ',')

# options(digits = 10)
# params %>% group_by(tc, harvConst) %>% summarise(min(rVal))

## some straight values for manuscript
# median(test$lifehist[,4:7]) ## median Omega tau (fecundity vector), mature sizes only
# sd(test$lifehist[,4:7]) ## sd Omega tau (fecundity vector)
# c(quantile(test$lifehist[,4:7],.025,names=FALSE),quantile(test$lifehist[,4:7],.975,names=FALSE))


# median(test$lifehist[,8:13]) ## median Omega tau (surv vector)
# sd(test$lifehist[,8:13]) ## median Omega tau (surv vector)
# hist(test$lifehist[,8:13])

# with(subset(test$outUs, tc == 0), min(Us))
# with(subset(test$outUs, tc == 0), sd(Us))
 
## Survivorship parameters 
# as.data.frame(test$lifehist[,9:15]) %>% summarise_each(funs(mean))
# as.data.frame(test$lifehist[,9:15]) %>% summarise_each(funs(median))
# as.data.frame(test$lifehist[,9:15]) %>% summarise_each(funs(quantile,0.25))
# quantile(test$lifehist[,9:15],.025,names=FALSE)
 
```



